LEGWORK: A python package for computing the evolution and detectability \\ of stellar-origin gravitational-wave sources with space-based detectors

### 2022-01-29 src/ms.tex
re-organise strain section

It is often convenient to convert between orbital frequency, $f_{\rm orb}$, and the semi-major axis, $a$, of a binary and this can be accomplished with Kepler's third law.
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    a = \qty(\frac{G(m_1 + m_2)}{(2 \pi f_{\rm orb})^{2}})^{1/3},
    \label{eq:kepler3rd}
\end{equation}
where $G$ is the gravitational constant. Inversely,
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    f_{\rm orb} = \frac{1}{2 \pi} \sqrt{\frac{G(m_1 + m_2)}{a^3}}.
    \label{eq:kepler3rd_reverse}
\end{equation}
For circular binaries, gravitational-wave emission occurs at twice the orbital frequency ($f_{\rm GW} = 2 f_{\rm orb}$). However, for eccentric binaries, we need to consider all frequency harmonics of gravitational-wave emission. These are defined such that the $n^{\rm th}$ harmonic frequency is
\begin{equation}{}
    f_{n} \equiv n \cdot f_{\rm orb}.
    \label{eq:nth_freq}
\end{equation}
It will be important to know the relative gravitational-wave power radiated into the $n^{\rm th}$ harmonic for a binary with eccentricity $e$ for the following strain and SNR calculations. This is given by \citet[][Eq.\,20]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_g}}
    \begin{aligned}
        g(n, e) = \frac{n^{4}}{32} & \Big\{ \Big[ J_{n-2}(n e)-2 e J_{n-1}(n e)+\frac{2}{n} J_{n}(n e) \\
        &\left. +2 e J_{n+1}(n e)-J_{n+2}(n e)\Big]^{2}\right.\\
        &\left.\kern-\nulldelimiterspace\right.
            \!\!\begin{aligned}
            +\big(1-e^{2}\big)\Big[&J_{n-2}(n e)-2 J_{n}(n e) \\
            &+J_{n+2}(n e)\Big]^{2}
            \end{aligned}\\
        &+\frac{4}{3 n^{2}}\Big[J_{n}(n e)\Big]^{2}\Big\},
    \end{aligned}
    \label{eq:g(n,e)}
\end{equation}
where $J_{n}(v)$ is the Bessel function of the first kind. Thus, the \textit{sum} of $g(n, e)$ over all harmonics gives the factor by which the gravitational-wave emission is stronger for a binary of eccentricity $e$ over an equivalent circular binary. This enhancement factor is \citep[][Eq.\,17]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_f}}
    F(e) \equiv \sum_{n = 1}^{\infty} g(n, e) = \frac{1 + \frac{73}{24} e^2 + \frac{37}{96} e^4}{(1 - e^2)^{7/2}}.
    \label{eq:eccentricity_enhancement_factor}
\end{equation}
A useful rule of thumb is that $F(0.5) \approx 5.0$, or in words, a binary with eccentricity $0.5$ loses energy to gravitational waves at 5 times the rate of an equivalent circular binary.

@@

\subsection{Strains {\normalfont (\texorpdfstring{\lwModLink{strain}}{strain})}}\label{sec:deriv_strain}
The strength of a gravitational wave in a detector at any one moment is determined by the strain amplitude, $h_0$. However, for stellar-origin sources at mHz frequencies, the signal can be present in the detector for many years. This means that, the $n^{\rm th}$ harmonic of the binary will spend approximately $f_n / \dot{f}_n$ seconds (or $f_n^2 / \dot{f}_n$ cycles) in the vicinity of a frequency $f_n$ \citep{Finn+2000}. This leads to the signal `accumulating' at the frequency $f_n$.

Therefore, to account for the integration of the signal over the mission, we instead use the `characteristic' strain amplitude of the $n^{\rm th}$ harmonic \citep[e.g][]{Finn+2000, Moore+2015}\footnote{Note that this is factor of 2 different from \citet{Finn+2000}. This is because the factor of 2 is already included in the \citet{Robson+2019} sensitivity curve and so is removed here.}
\begin{equation}{}
    h_{c, n}^2 = \qty(\frac{f_n^2}{\dot{f}_n}) h_{0, n}^2,
    \label{eq:strain-charstrain}
\end{equation}
where $h_{0, n}$ is the strain amplitude in the $n^{\rm th}$ harmonic.

The characteristic strain represents the strain measured by the detector over the duration of the mission (approximated as a single broad-band burst), whilst the strain amplitude is the strength of the GW emission at each instantaneous moment.

\subsubsection{Characteristic Strain}

The characteristic strain from a stellar mass binary in the $n^{\rm th}$ harmonic is given by (e.g. \citealp[][Eq.\,56]{Barack+2004}; \citealp[][Eq.\,5.1]{ Flanagan+1998})
\begin{equation}{}
    h_{c,n}^2 = \frac{1}{(\pi D_L)^2} \qty( \frac{2 G}{c^3} \frac{\dot{E_n}}{\dot{f_n}} ),
    \label{eq:char_strain_dedf}
\end{equation}
where $D_L$ is the luminosity distance to the source (note that for Milky Way sources, or any sources with redshift $\sim 0$, this is simply the distance to the source), $\dot{E}_n$ is the power radiated in the $n^{\rm th}$ harmonic and $\dot{f}_n$ is the rate of change of the $n^{\rm th}$ harmonic frequency.

>>>>

It is often convenient to convert between orbital frequency, $f_{\rm orb}$, and the semi-major axis, $a$, of a binary and this can be accomplished with Kepler's third law.
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    a = \qty(\frac{G(m_1 + m_2)}{(2 \pi f_{\rm orb})^{2}})^{1/3},
    \label{eq:kepler3rd}
\end{equation}
where $G$ is the gravitational constant. Inversely,
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    f_{\rm orb} = \frac{1}{2 \pi} \sqrt{\frac{G(m_1 + m_2)}{a^3}}.
    \label{eq:kepler3rd_reverse}
\end{equation}
For circular binaries, gravitational-wave emission occurs at twice the orbital frequency ($f_{\rm GW} = 2 f_{\rm orb}$). However, for eccentric binaries, we need to consider all frequency harmonics of gravitational-wave emission. These are defined such that the $n^{\rm th}$ harmonic frequency is
\begin{equation}{}
    f_{n} \equiv n \cdot f_{\rm orb}.
    \label{eq:nth_freq}
\end{equation}
It will be important to know the relative gravitational-wave power radiated into the $n^{\rm th}$ harmonic for a binary with eccentricity $e$ for the strain and SNR calculations. This is given by \citet[][Eq.\,20]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_g}}
    \begin{aligned}
        g(n, e) = \frac{n^{4}}{32} & \Big\{ \Big[ J_{n-2}(n e)-2 e J_{n-1}(n e)+\frac{2}{n} J_{n}(n e) \\
        &\left. +2 e J_{n+1}(n e)-J_{n+2}(n e)\Big]^{2}\right.\\
        &\left.\kern-\nulldelimiterspace\right.
            \!\!\begin{aligned}
            +\big(1-e^{2}\big)\Big[&J_{n-2}(n e)-2 J_{n}(n e) \\
            &+J_{n+2}(n e)\Big]^{2}
            \end{aligned}\\
        &+\frac{4}{3 n^{2}}\Big[J_{n}(n e)\Big]^{2}\Big\},
    \end{aligned}
    \label{eq:g(n,e)}
\end{equation}
where $J_{n}(v)$ is the Bessel function of the first kind. Thus, the \textit{sum} of $g(n, e)$ over all harmonics gives the factor by which the gravitational-wave emission is stronger for a binary of eccentricity $e$ over an equivalent circular binary. This enhancement factor is \citep[][Eq.\,17]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_f}}
    F(e) \equiv \sum_{n = 1}^{\infty} g(n, e) = \frac{1 + \frac{73}{24} e^2 + \frac{37}{96} e^4}{(1 - e^2)^{7/2}}.
    \label{eq:eccentricity_enhancement_factor}
\end{equation}
A useful rule of thumb is that $F(0.5) \approx 5.0$, or in words, a binary with eccentricity $0.5$ loses energy to gravitational waves at 5 times the rate of an equivalent circular binary.

@@

\subsection{Strains {\normalfont (\texorpdfstring{\lwModLink{strain}}{strain})}}\label{sec:deriv_strain}
\referee{\subsubsection{Characteristic Strain}
The strength of a gravitational wave in a detector at any one moment is determined by the strain amplitude, $h_0$. However, for stellar-origin sources at mHz frequencies, the signal can be present in the detector for many years. This means that, the $n^{\rm th}$ harmonic of the binary will spend approximately $f_n / \dot{f}_n$ seconds (or $f_n^2 / \dot{f}_n$ cycles) in the vicinity of a frequency $f_n$ \citep{Finn+2000}. This leads to the signal `accumulating' at the frequency $f_n$.

Therefore, to account for the integration of the signal over the mission, we instead use the `characteristic' strain amplitude of the $n^{\rm th}$ harmonic, $h_{c, n}$, which is the term present in the general signal-to-noise ratio equation. This can be related the to the strain amplitude in the $n^{\rm th}$ harmonic, $h_{0, n}$, as \citep[e.g][]{Finn+2000, Moore+2015}\footnote{Note that this is factor of 2 different from \citet{Finn+2000}. This is because the factor of 2 is already included in the \citet{Robson+2019} sensitivity curve and so is removed here.}
\begin{equation}{}
    h_{c, n}^2 = \qty(\frac{f_n^2}{\dot{f}_n}) h_{0, n}^2,
    \label{eq:strain-charstrain}
\end{equation}
The characteristic strain represents the strain measured by the detector over the duration of the mission (approximated as a single broad-band burst), whilst the strain amplitude is the strength of the GW emission at each instantaneous moment. For a stellar mass binary, the characteristic strain in the $n^{\rm th}$ harmonic is given by (e.g. \citealp[][Eq.\,56]{Barack+2004}; \citealp[][Eq.\,5.1]{ Flanagan+1998})}
\begin{equation}{}
    h_{c,n}^2 = \frac{1}{(\pi D_L)^2} \qty( \frac{2 G}{c^3} \frac{\dot{E_n}}{\dot{f_n}} ),
    \label{eq:char_strain_dedf}
\end{equation}
where $D_L$ is the luminosity distance to the source (note that for Milky Way sources, or any sources with redshift $\sim 0$, this is simply the distance to the source), $\dot{E}_n$ is the power radiated in the $n^{\rm th}$ harmonic and $\dot{f}_n$ is the rate of change of the $n^{\rm th}$ harmonic frequency.
